A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-11-04 latex/main.tex
Clear up details on Pango X lineage events

\subsection{Pango X lineage events}
\label{sec:supplement_x_pango_lineages}
[TODO DOCUMENT DETAILS of deriving events]

The nested samples not assigned to the focal pango lineage are
either from different pango X lineages (see next paragraph), 
or given a generic BA.2 assignment. Of these 59 generic assignents
6 had conflicts in their UShER placements and could have been
assigned an X lineage (XJ: 3, XU: 2, XN: 1), and 
25 were cases in which Scorpio~\cite{Colquhoun2023scorpio}
failed to find a specific lineage (accounting for 7\%
of the 357 ``Omicron (Unassigned)`` Scorpio calls across
2,747,985 alignments).

>>>>

\subsection{Pango X lineage events}
\label{sec:supplement_x_pango_lineages}
Most of the Pango X lineages had straightforward origination events 
as defined in Section~\ref{sec:supplement_pango_lineages} 
and are represented in Table~\ref{tab:pango_x_lineages}.
Three lineages were treated specially, 
corresponding to 23 omitted samples in total.

XM has 4 separate origins, 3 singletons and one dominant node
with 26 descendants. We include the dominant node in 
Table~\ref{tab:pango_x_lineages} and omit the singletons. See
Document~\ref{sec:dd_xm} for further details.
XAC has 4 separate origins, and has been omitted from 
Table~\ref{tab:pango_x_lineages}. However, as Figure~\ref{fig:subgraphs}B
shows, these are all descending from a single BA.2 classified node 
descending from the putative Xx event. 
Similarly, XAD has two independent origins descending from the Xx
event, and we omit it for simplicity.
See Document~\ref{sec:dd_xx} for more details on the Xx event.

In cases where nested samples not assigned to the focal pango lineage are
either from different pango X lineages 
or given a generic BA.2 assignment. Of these 59 generic assignents
6 had conflicts in their UShER placements and could have been
assigned an X lineage (XJ: 3, XU: 2, XN: 1), and 
25 were cases in which Scorpio~\cite{Colquhoun2023scorpio}
failed to find a specific lineage (accounting for 7\%
of the 357 ``Omicron (Unassigned)`` Scorpio calls across
2,747,985 alignments).

See notebook
\href{https://github.com/jeromekelleher/sc2ts-paper/blob/main/notebooks/tab\_pango\_x\_events.ipynb}
{tab\_pango\_x\_events.ipynb} for details.
